---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data

```{r load_data, echo=TRUE}

library(data.table)
library(ggplot2)

df <- read.csv('./activity.csv')
DT <- data.table(df)
summary(DT)
str(DT)

```
## What is mean total number of steps taken per day?
Total steps by day
```{r, echo=TRUE}
DT[, sum(steps), by=date]$V1
```

Mean and median of steps per day
```{r mean_steps, echo=TRUE}

DT[, list(as.double(median(steps)), mean(steps)), by=date]

```

## What is the average daily activity pattern?


Histrogram of steps by day
```{r hist_steps, echo=TRUE}

hist(x=DT[, sum(steps), by=date]$V1, xlab = "Steps by date", main = "Histogram fo Steps by Date")


```

Activity pattern

```{r plot_pattern, echo=TRUE}

plot(DT[, mean(steps, na.rm = TRUE), by=date]$V1, ylab = "Average steps", 
     xlab = "Day", main = "Average steps per day plot", type = "l")

```

Interval with maximum steps
```{r max_interval, echo=TRUE}
DT[steps == max(DT$steps),interval]

```

## Imputing missing values

Number of missing values 
```{r NA_qty}
# 1 number of NAs
DT[is.na(steps),.N]
```

Removing NA values
```{r del_NA, echo=TRUE}
DT2 <- DT[!is.na(steps)]

```


Histograms, mean and median without missing values
```{r hist_without_NA, echo=TRUE}
# histogram, mean and median without missing values
hist(x=DT2[, sum(steps), by=date]$V1, xlab = "Steps by date", main = "Histogram fo Steps by Date whitout missing values")

DT2[, list(as.double(median(steps)), mean(steps)), by=date]

```

## Are there differences in activity patterns between weekdays and weekends?

Create a factor column with levels; weekend and weekday
Note that compare day names in spanish due to my computer configuration
```{r typeday_column, echo=TRUE}
DT2[, day.type := "weekday"]
DT2[weekdays(as.Date(date)) %in% c("sabado", "domingo"), day.type := "weekend"]
DT2$day.type <- factor(DT2$day.type)

```

Plotting comparing setps patterns between weekday and weekend
```{r plot_by_typeday, echo=TRUE}
par(mfrow=c(2,1))
plot(DT2[day.type == "weekend", mean(steps), by = interval], type="l", main="Average steps by interval in weekend", ylab="average steps")
plot(DT2[day.type == "weekday", mean(steps), by = interval], type="l", main="Average steps by interval in weekday", ylab="average steps")

```
